x402 and related programmable-payment protocols make it increasingly practical for autonomous agents to pay for online services without an interactive checkout step, but supply no algorithm for the buyer's first decision: which provider to pay. We formalize this gap as \emph{agent-native payment decision-making}---budget-aware, non-stationary, contextual provider selection over heterogeneous paid services, where the buyer is an autonomous wallet-bearing agent rather than a human-operated routing service. We propose \emph{402Pilot}, a protocol-agnostic decision layer with stable interfaces between agent decision-making and payment execution, and \emph{PA-DCT} (Payment-Aware Discounted Contextual Thompson sampling), the learning policy that drives it. PA-DCT maintains discounted dual Gaussian posteriors over both quality and cost per (provider, task-type) bucket, treating cost as a learned signal rather than a known scalar so that price shocks are detected the same way quality shocks are. We construct \emph{402Pilot-Bench}, a reproducible replay benchmark of $823$ tasks across three domains (coding, multi-hop QA, web search) split into four evaluation buckets (web search separates closed-form and open-ended due to different evaluators); five heterogeneous LLM providers including adversarial and flaky variants at the same price tier; three calibrated market scenarios (stationary, mid-tier outage, premium price promotion); and $30$ seeds $\times$ $10{,}000$ rounds per cell, totaling $20{,}575$ pre-generated frozen LLM responses. In the non-stationary scenarios, PA-DCT wins $13$ of $14$ baseline comparisons on cumulative payment-aware reward; the remaining comparison is a statistically non-significant tie against Always-P-cheap in the mid-tier outage, where PA-DCT delivers materially higher mean quality. It incurs a measured exploration cost in the stationary baseline, where adaptation provides no benefit. An ablation over PA-DCT's four components (payment-awareness, discount, contextual bucketing, Thompson sampling) plus a diagnostic isolation of the learned cost posterior inside payment-awareness shows that no single classical metric reveals all of these contributions; we report all five economic metrics (cum-PA, ROI, mean quality, regret, adaptation time) and recommend the multi-metric framework for future work in this setting.
