\section{Related Work}
\label{sec:related-work}

\paragraph{Agent payment protocols.}
Recent agent-payment protocols make it possible for software agents to pay for services at machine speed. x402 embeds per-request payment into HTTP and its V2 release adds automatic discovery, dynamic payment recipients, extensible payment schemes, and dynamic service metadata~\citep{coinbase2025x402v2}. AP2 provides a payment-agnostic authorization framework for agent-led commerce~\citep{google2025ap2}, while A2A--x402 brings x402-style cryptocurrency payments into the Agent-to-Agent protocol~\citep{a2ax402}. A402 targets a different layer: it binds cryptocurrency payment to service execution using atomic service channels~\citep{a402}. A recent SoK classifies blockchain-based agent payments around discovery, authorization, execution, and accounting~\citep{sok2026agentpayments}. These works answer how an agent can authenticate, authorize, and settle a payment. They do not answer which provider an autonomous agent should pay under a finite wallet, uncertain service quality, and changing prices. 402Pilot is therefore not a fork or replacement of x402: it is a decision layer that sits before payment execution and can, in principle, be attached to x402 or future programmable payment protocols.

\paragraph{Discovery, ranking, and the missing buyer-side policy.}
The x402 V2 Discovery extension makes candidate services visible through structured metadata and facilitator-side indexing~\citep{coinbase2025x402v2}. This turns endpoint selection into an explicit buyer-side problem: an agent may see multiple payable services with different prices, failure rates, and historical outcomes. Existing protocol documents and ecosystem registries expose the information needed for ranking, but leave the ranking policy itself unspecified. 402Pilot fills this algorithmic gap by treating provider choice as an online decision problem: after each paid call, observed utility, failure, and realized cost are fed back into the selector. This separation is important. Protocols provide the settlement substrate; 402Pilot provides a protocol-agnostic policy interface for deciding what to buy next.

\paragraph{Budgeted, contextual, and non-stationary bandits.}
Classical bandit work provides several ingredients for online provider selection under uncertainty. Budgeted Thompson Sampling samples both reward and cost posteriors and selects arms by reward--cost tradeoff under a finite budget~\citep{xia2015bts}; bandits with knapsacks provide the broader constrained-resource framing~\citep{badanidiyuru2018bandits}. Discounted Thompson Sampling addresses non-stationary rewards by exponentially down-weighting old observations~\citep{qi2023dsts}. Contextual Thompson Sampling models action value as a function of task context~\citep{agrawal2013contextualts}, with LinUCB as a canonical frequentist counterpart~\citep{li2010linucb}. 402Pilot differs from this literature in its operating regime rather than by treating any one ingredient as new: the learner is the spending agent itself, payments are made per call, costs are observed through market-facing quotes, and the policy must adapt without a human operator resetting budgets or intervening after price and quality shocks.

\paragraph{Cost-aware LLM routing and adaptive selection.}
A growing literature studies quality--cost tradeoffs in LLM routing. FrugalGPT uses cascades to reduce cost while preserving answer quality~\citep{chen2023frugalgpt}; RouteLLM, Hybrid LLM, AutoMix, MixLLM, and LLM Bandit learn to assign prompts to cheaper or stronger models based on predicted quality, preference, cost, or latency~\citep{ong2024routellm,ding2024hybridllm,madaan2024automix,mixllm2024,llmbandit2025}. These systems are close in motivation, but they generally assume a platform-controlled router, published price cards, and offline or semi-online evaluation over static model APIs.

\paragraph{Reverse auctions and mechanism design over agent markets.}
Truthful reverse auctions for adaptive selection via contextual multi-armed bandits study an adjacent provider-selection problem in which strategic providers submit bids and the mechanism designer seeks truthful allocation and payment rules~\citep{aamas2026reverseauction}. 402Pilot studies a different market interface: x402-style quote-and-pay APIs, where the buyer observes payable endpoints and must adapt its own spending policy over time. Thus, prior LLM routing and reverse-auction mechanisms address model selection and incentive compatibility; 402Pilot addresses autonomous buyer-side adaptation under wallet pressure and non-stationary provider behavior.

\paragraph{Positioning against prior bandit and routing methods.}
Table~\ref{tab:prior-bandits} summarizes how representative prior methods cover the three agent-native ingredients (budget, drift, context) and the underlying setting. Each prior method addresses a subset; PA-DCT combines all three in the agent-native micropayment-buyer regime.

\begin{table}[t]
\centering
\caption{Coverage of the three agent-native ingredients across representative bandit and LLM-routing methods. ``Setting'' marks methods designed (rather than retrofitted) for an autonomous buyer over programmable micropayments. PA-DCT is the only entry that combines all four.}
\label{tab:prior-bandits}
\begin{tabular}{lcccc}
\toprule
Method & Budget & Discount & Context & Setting \\
\midrule
BTS~\citep{xia2015bts}                          & \checkmark & --         & --         & -- \\
DS-TS~\citep{qi2023dsts}                        & --         & \checkmark & --         & -- \\
Contextual TS~\citep{agrawal2013contextualts}   & --         & --         & \checkmark & -- \\
Sliding-Window TS                               & --         & \checkmark & --         & -- \\
TV-Reward Contextual TS                         & --         & \checkmark & \checkmark & -- \\
C2MAB-V (LLM routing)                           & partial    & --         & \checkmark & -- \\
NeuralLinUCB~\citep{li2010linucb}               & --         & --         & \checkmark & -- \\
Combinatorial + change-point                    & \checkmark & \checkmark & --         & -- \\
\midrule
\textbf{PA-DCT (ours)}                          & \checkmark & \checkmark & \checkmark & \checkmark \\
\bottomrule
\end{tabular}
\end{table}
