\section{Problem Formulation}
\label{sec:problem}

We formalize agent-native payment decision-making as a budgeted, non-stationary contextual bandit over protocol-quoted payable endpoints. The core loop is quote--pay--observe: a buyer sees a task and payment quotes, pays exactly one affordable provider before seeing service quality, and updates from only that paid call.

\paragraph{Round protocol.}
Consider a single autonomous buyer with initial wallet budget $B$ and $K$ payable service providers over rounds $t \in \{0,\ldots,T-1\}$. At the start of round $t$, the buyer observes a task context $\mathbf{x}_t \in \mathcal{X}$ and the wallet state $B_t$. A programmable payment interface exposes enough price information to determine which providers are affordable before payment; let $p_{a,t}$ denote the pre-payment quote for provider $a$, write $\mathbf{p}_t=(p_{1,t},\ldots,p_{K,t})$, and define
\begin{equation}
\label{eq:affordable-set}
\mathcal{A}_t = \{a \in \{1,\ldots,K\}: p_{a,t} \leq B_t\}.
\end{equation}
If $\mathcal{A}_t$ is empty, the run stops. Otherwise, the buyer selects one provider $a_t \in \mathcal{A}_t$, commits to the paid call through the protocol, and cannot observe the service outcome before paying. The selected call then produces the only feedback sample for the round:
\[
(q_t, c_t, f_t) \sim \mathcal{D}_{a_t,t}(\cdot \mid \mathbf{x}_t),
\]
where $q_t \in [0,1]$ is a task-level quality score, $c_t \in \mathbb{R}_{\geq 0}$ is the realized charged cost, and $f_t \in \{0,1\}$ marks timeout, parse failure, or another non-delivery event. The quote $p_{a,t}$ determines affordability before selection; the realized charge $c_t$ is logged after the selected payment. In fixed quoted-price calls these coincide, $c_t=p_{a_t,t}$, while the separate notation allows discounts, fees, or post-quote adjustments. Outcomes and realized charges for unchosen providers are not observed, which makes provider choice a bandit problem rather than a full-information ranking problem. The wallet then evolves as
\begin{equation}
\label{eq:wallet}
B_{t+1} = B_t - c_t, \qquad B_0 = B
\end{equation}
The stopping time $\tau$ equals $T$ if the run reaches the horizon, and otherwise is the first round with $\mathcal{A}_t=\emptyset$.

The environment is the sequence of tasks, quotes, and provider outcome laws
\[
\mathcal{E}=\{(\mathbf{x}_t,\mathbf{p}_t,\mathcal{D}_{1,t},\ldots,\mathcal{D}_{K,t})\}_{t=0}^{T-1}.
\]

\paragraph{Admissible policies.}
Let $\mathcal{H}_t$ denote the information available immediately before selecting $a_t$:
\[
\mathcal{H}_t =
\sigma\!\left(
\{\mathbf{x}_s,\mathbf{p}_s,a_s,q_s,c_s,f_s\}_{s<t},
\mathbf{x}_t,\mathbf{p}_t,B_t
\right).
\]
An admissible policy is a non-anticipating sequence $\pi=\{\pi_t\}_{t=0}^{T-1}$ satisfying
\[
a_t \sim \pi_t(\cdot \mid \mathcal{H}_t),
\qquad
\mathrm{supp}(\pi_t) \subseteq \mathcal{A}_t .
\]
Thus the policy may use the current task, wallet state, available quotes, and past paid outcomes, but not unchosen outcomes, future quotes, or future shocks.

\paragraph{Payment-aware objective.}
The round feedback provides two signals: service delivery and realized payment cost. We first summarize service delivery as utility,
\begin{equation}
\label{eq:utility}
\ut = q_t - \nu f_t.
\end{equation}
The failure penalty captures retry overhead, broken call chains, and downstream agent stalls that are not fully represented by setting $q_t=0$ on failure. We then combine utility with normalized realized cost to obtain the payment-aware reward,
\begin{equation}
\label{eq:reward}
\ctilde = \mathrm{clip}(c_t/c_{\max},0,1), \qquad
\lamn = \frac{\lambda_t}{1+\lambda_t}, \qquad
\rt = (1-\lamn)\ut - \lamn\ctilde .
\end{equation}
Here $\mathrm{clip}(z,0,1)=\min\{1,\max\{0,z\}\}$, $c_{\max}>0$ is a fixed cost normalizer chosen before evaluation, and $\lambda_t>0$ is the wallet-pressure multiplier defined below. The bounded weight $\lamn \in (0,1)$ keeps utility and cost comparable across wallet states while preserving the sign of failures. For $0\leq\nu\leq1$, $\ut\in[-1,1]$; since $\ctilde\in[0,1]$, the payment-aware reward satisfies $\rt\in[-1,1]$. Under environment $\mathcal{E}$, a policy $\pi$ aims to maximize
\begin{equation}
\label{eq:objective}
J_{\mathcal{E}}(\pi) = \mathbb{E}^{\pi}_{\mathcal{E}}\!\left[\sum_{t=0}^{\tau-1} \rt \right],
\end{equation}
where the expectation is over policy randomization and selected-provider outcomes. We write $\Reg_{\mathcal{E}}(\pi;\bar\pi)=J_{\mathcal{E}}(\bar\pi)-J_{\mathcal{E}}(\pi)$ for regret against a comparator $\bar\pi$. In experiments, $\bar\pi$ is the full-information True Oracle of \S\ref{sec:eval-comparators}; because it observes unchosen outcomes, it is an empirical upper bound rather than an admissible online policy. In PA-DCT, beliefs are updated from the two observed channels $(\ut,c_t)$ rather than from the scalar $\rt$: $\ut$ tracks provider quality, $c_t$ tracks realized payment cost, and $\rt$ defines the decision/evaluation objective.

\paragraph{Budget pressure.}
It remains to define the wallet-pressure multiplier $\lambda_t$ used in Eq.~\eqref{eq:reward}. This multiplier is a deterministic function of cumulative spend. For $t>0$, define the realized burn rate as the fraction of the wallet spent per elapsed round,
\begin{equation}
\label{eq:burn-rate}
\mathrm{actual\_burn\_rate}_t = \frac{(B-B_t)/B}{t},
\qquad
\mathrm{target\_rate} = \frac{1}{T}
\qquad (t>0).
\end{equation}
The normalized burn deviation is
\begin{equation}
\label{eq:lambda}
\bdev_t =
\frac{\mathrm{actual\_burn\_rate}_t-\mathrm{target\_rate}}
{\mathrm{target\_rate}},
\qquad
\lambda_t =
\begin{cases}
1, & t=0,\\
\exp(\alpha \bdev_t), & t>0,
\end{cases}
\end{equation}
with sharpness parameter $\alpha \geq 0$. Thus $\lamn=0.5$ when the buyer is exactly on the linear burn plan, decreases when the buyer is under-spending, and increases toward cost-dominated decisions when the wallet is being spent too quickly.

\paragraph{Non-stationarity and learning class.}
We allow both $\mathcal{D}_{a,t}(\cdot \mid \mathbf{x})$ and $p_{a,t}$ to vary with $t$, so quality, reliability, and effective price may drift. Change points are not announced to the policy. Tasks arrive from an exogenous workload stream, and providers are non-strategic posted-price endpoints: they expose quotes and deliver services rather than bidding strategically for allocation. Reverse auctions, multi-agent coordination, dispute resolution, and latency-aware routing are outside this formulation.
Within this scope, the only controlled action is the provider choice $a_t$; the immediate feedback is the selected call's $(\ut,c_t)$. Apart from wallet accounting and the policy's internal belief state, actions do not induce latent environment transitions. The policy must adapt to drift, but it does not control or model a Markovian transition process over latent states. We therefore model agent-native payment decision-making as a contextual bandit with budget state and non-stationary feedback.
