\section{The 402Pilot Decision Layer}
\label{sec:layer}

402Pilot is a decision layer between an autonomous agent and a programmable payment protocol such as x402. The layer accepts a task and a wallet state, picks a provider, calls the protocol stack to settle the payment, and feeds the realized outcome back into a policy. It does not modify the protocol, never holds funds, and never crosses the settlement boundary. Figure~\ref{fig:architecture} sketches the per-round flow.

\begin{figure}[t]
\centering
\resizebox{0.98\textwidth}{!}{%
\begin{tikzpicture}[
  font=\footnotesize,
  block/.style={
    draw=black!55, rounded corners=2pt, align=center,
    minimum height=0.95cm, text width=2.2cm, inner sep=4pt, fill=white
  },
  pilotblock/.style={block, draw=blue!55!black, fill=blue!5},
  extblock/.style={block, draw=black!45, fill=black!7},
  payblock/.style={block, draw=black!45, fill=black!10, dashed},
  ifblock/.style={block, draw=blue!55!black, fill=white, line width=0.7pt, text width=2.0cm},
  arrow/.style={-{Stealth[length=2.0mm]}, line width=0.5pt},
  feedback/.style={-{Stealth[length=2.0mm]}, line width=0.5pt, dashed},
  lab/.style={font=\scriptsize, fill=white, inner sep=1.5pt, align=center}
]

% --- Node placement (manual coordinates, units in cm) ---
% Top row at y=2.0 (forward path); bottom row at y=0 (state + feedback)
\node[extblock] (agent) at (-1.0, 2.0) {Agent};
\node[pilotblock] (ctx) at (3.0, 2.0) {Context\\encoder\\$\mathbf{x}_t$};
\node[pilotblock, text width=2.4cm] (select) at (7.0, 2.0) {PA-DCT\\\texttt{policy.select}\\$\to a_t$};
\node[payblock] (x402) at (13.0, 2.0) {x402\\stack};
\node[payblock] (provider) at (16.5, 2.0) {Provider};

\node[pilotblock] (wallet) at (3.0, 0) {Wallet\\$\lambda_t,\,\mathcal{A}_t$};
\node[pilotblock, text width=2.4cm] (update) at (7.0, 0) {\texttt{policy.update}\\$u_t = q_t - \nu f_t$};
% Outcome scorer: center sits ON the layer boundary, so left half ends up inside
% pilotbox and right half inside paybox (both fits include scorer.center).
\node[ifblock] (scorer) at (11.0, 0) {Outcome\\scorer\\$(q_t,c_t,f_t)$};

% --- Background boxes ---
% Trick: define two coords offset by inner_xsep on either side of scorer.center.
% After fit's inner_xsep is added, pilotbox right edge and paybox left edge both
% land EXACTLY on scorer.center — no overlap, no gap. The two boxes share a
% vertical line that bisects the Outcome scorer.
\coordinate (bndL) at ([xshift=-8pt]scorer.center);
\coordinate (bndR) at ([xshift=8pt]scorer.center);

\begin{scope}[on background layer]
  \node[
    draw=blue!45!black, fill=blue!4,
    rounded corners=4pt, inner xsep=8pt, inner ysep=14pt,
    fit=(ctx) (select) (wallet) (update) (bndL)
  ] (pilotbox) {};
  \node[
    draw=black!35, fill=black!5,
    rounded corners=4pt, dashed, inner xsep=8pt, inner ysep=10pt,
    fit=(x402) (provider) (bndR)
  ] (paybox) {};
\end{scope}

\node[anchor=north west, font=\scriptsize\bfseries, text=blue!55!black]
  at ([xshift=4pt,yshift=-4pt]pilotbox.north west) {402Pilot decision layer};
% paybox label parked at the bottom-right corner so it doesn't fight with arrow (5)
\node[anchor=south east, font=\scriptsize\bfseries, text=black!55]
  at ([xshift=-4pt,yshift=4pt]paybox.south east) {protocol stack + provider};

% --- Forward arrows (top row) ---
\draw[arrow] (agent) -- node[lab, above] {(1) task} (ctx);
\draw[arrow] (ctx) -- node[lab, above] {(2) context\\$\mathbf{x}_t$} (select);
\draw[arrow] (select) -- node[lab, above] {(4) chosen arm\\$a_t$} (x402);
\draw[arrow] (x402) -- node[lab, above] {(5) pay\\+ call} (provider);

% Wallet → select (diagonal up-right). Label kept compact because the diagonal
% is short; the caption carries the prose "budget pressure + affordable set".
\draw[arrow] (wallet.north east) -- node[lab, sloped, pos=0.55]
  {(3) $\lambda_t,\,\mathcal{A}_t$} ([xshift=-6pt]select.south);

% Provider → scorer (response crosses external space, lands at the straddling node)
\draw[arrow] (provider.south) -- node[lab, sloped, pos=0.55] {(6) response} (scorer.north east);

% --- Feedback arrows (bottom row) ---
% Scorer → update: outcome tuple flows back into the policy
\draw[arrow] (scorer) -- node[lab, above] {(7) outcome\\$(q_t,c_t,f_t)$} (update);

% Update → select: belief update for next round (slight x-shift to avoid wallet→select)
\draw[feedback] ([xshift=8pt]update.north) -- node[lab, right, pos=0.5]
  {(8) update\\$u_t,\, c_t$} ([xshift=8pt]select.south);

% Update → wallet (record realized spend)
\draw[feedback] (update.west) -- node[lab, above] {\texttt{record\_spend}} (wallet.east);

\end{tikzpicture}%
}
\caption{The 402Pilot decision-layer architecture. The Outcome scorer straddles the layer boundary by design: the right half receives raw responses from the (external) provider; the left half emits the structured outcome $(q_t,c_t,f_t)$ to \texttt{policy.update} (inside the layer). Per round: (1) the agent submits a task; (2) the context encoder produces $\mathbf{x}_t$; (3) the wallet exposes budget pressure $\lambda_t$ and the affordable set $\mathcal{A}_t$; (4) \texttt{policy.select} returns chosen arm $a_t$; (5) the x402 stack settles and calls the chosen provider (gray dashed: outside the layer); (6) the provider's response is handed to the Outcome scorer; (7) the scorer emits $(q_t,c_t,f_t)$; (8) \texttt{policy.update} incorporates utility $u_t = q_t - \nu f_t$ and observed cost $c_t$ --- \emph{not} the payment-aware reward $r_t$. Settlement is delegated unchanged to the protocol stack.}
\label{fig:architecture}
\end{figure}

\paragraph{Two stable interfaces.}
A policy is any object exposing
\begin{align*}
&\mathtt{select}(\mathbf{x}_t,\, \mathcal{A}_t,\, \mathrm{wallet\_state}) \to a_t \in \mathcal{A}_t, \\
&\mathtt{update}(\mathbf{x}_t,\, a_t,\, \ut,\, c_t) \to \emptyset,
\end{align*}
where $\mathcal{A}_t \subseteq \{1, \ldots, K\}$ is the set of providers that fit in the remaining wallet at the current effective prices. Affordability is enforced by the wallet, not the policy: a policy never has to reason about hard budget constraints, only about picking well from a feasible set. The two methods are the only points of contact with the rest of the stack; everything else (priors, posteriors, internal state) is private to the policy.

\paragraph{Update signal: utility, not payment-aware reward.}
A subtle but load-bearing commitment: the layer feeds \emph{utility} $\ut$ to \texttt{policy.update}, not the payment-aware reward $\rt$. This decoupling lets the posterior track \emph{intrinsic provider quality} independently of decision-time budget pressure. If we instead updated on $\rt$, beliefs about a provider would shift purely because the wallet drained, defeating the purpose of having a posterior. Concretely: the same provider should not look worse to the policy late in a budget-tight run than it did early in the run when the wallet was full. Decision-time budget pressure enters only inside \texttt{select}, through $\lamn$, never inside \texttt{update}.

\paragraph{Pluggability across policies and protocols.}
On the policy side, every comparator we evaluate (\S\ref{sec:eval-comparators})---Random, three Always-X variants, BudgetRule, and PA-DCT itself---implements the same two-method interface. New algorithms drop in without touching the rest of the stack. On the protocol side, the layer is agnostic to which programmable payment substrate sits below: today we instantiate above x402~\citep{coinbase2025x402v2}, but the same interfaces apply to AP2~\citep{google2025ap2}, A2A--x402~\citep{a2ax402}, A402~\citep{a402}, or any future quote-and-pay protocol. The protocol stack provides addressing, settlement, and audit trails; 402Pilot provides the decision policy.

\paragraph{What the layer does \emph{not} do.}
Five concerns are explicitly out of scope and are addressed in \S\ref{sec:discuss}:
on-chain settlement (the protocol's job);
capability discovery (handled by x402 V2's Discovery extension or equivalent registries);
production observability (operator dashboards, alerting);
multi-agent coordination (we operate at single-agent scope: one wallet, one policy, one workload stream);
and online safeguards such as drift detection or trust scoring (future work; we discuss directions in \S\ref{sec:discuss}).

\paragraph{What the layer assumes.}
402Pilot makes three assumptions about the surrounding stack. First, the protocol can quote a price for a paid call before the call is made, so the wallet can compute affordability and the policy can include cost in its decision rule. Second, outcomes are observable and scoreable shortly after the call, so the policy can update before the next round (we make this concrete in our benchmark via deterministic per-task evaluators or cached LLM-as-judge scores). Third, providers are non-strategic in the protocol-mechanism sense: they post prices and deliver services rather than playing auction games over bids. All three assumptions hold for x402 V2 and its current peers. Relaxing the third---strategic providers---is the regime studied by recent mechanism-design work~\citep{aamas2026reverseauction}; 402Pilot is complementary to that line, not a substitute.

\paragraph{Implementation footprint.}
The reference implementation is a Python package (\texttt{pilot402}) of roughly $4{,}900$ source lines, with the policy interface, the wallet, the context encoder, the reward calculator, and the per-seed runtime loop as separate modules; any one of them can be swapped without touching the others. Replays are deterministic given the seed and the frozen pre-generated dataset, so the full main sweep ($5$ baselines $\times$ $3$ scenarios $\times$ $30$ seeds $\times$ $10{,}000$ rounds) finishes within a few CPU-hours on a single laptop, with no GPU dependency at run time because all LLM responses are cached.
