\section{Method: PA-DCT}
\label{sec:method}

PA-DCT is an instance of contextual Thompson sampling adapted to the agent-native regime (\S\ref{sec:intro}). Each (arm, task-type-bucket) cell maintains \emph{two} independent Gaussian posteriors---one over utility, one over cost---both tracked through discounted sufficient statistics. At decision time, the policy samples both posteriors and ranks affordable arms by a budget-pressure-weighted score; on receipt of the realized outcome, only the chosen cell's posteriors are updated, after a global discount sweep over every cell.

\subsection{Dual Gaussian posteriors}
\label{sec:method-posteriors}

For each arm $a \in \{1,\ldots,K\}$ and bucket $k \in \{1,\ldots,K_b\}$ (where $K_b$ is the number of context buckets, default $4$, one per task type), we maintain two posteriors:
\begin{itemize}
\item \textbf{Q-posterior}, tracking the per-arm utility $\ut = q_t - \nu f_t$. Conjugate Normal-Normal with prior $\mathcal{N}(\mu_{0,q}, \sigma^2_{0,q})$ and likelihood variance $\sigma^2_q$.
\item \textbf{C-posterior}, tracking the per-arm raw cost $c_t$. Same Normal-Normal form, with prior $\mathcal{N}(\bar c_a, \sigma^2_{0,c})$ where $\bar c_a$ is the spec base price for arm $a$, and likelihood variance $\sigma^2_c$.
\end{itemize}

Each cell carries the standard discounted sufficient statistics $n^{a,k}_t \in \mathbb{R}_{\geq 0}$ (effective sample count) and $S^{a,k}_t \in \mathbb{R}$ (weighted sum of observations), with the analogous pair for the cost posterior. Posterior parameters follow the closed-form Normal-Normal expressions, e.g.\ for Q:
\begin{equation}
\label{eq:posterior-q}
\Sigmaq{a}{k} = \left(\frac{1}{\sigma^2_{0,q}} + \frac{n^{a,k}_t}{\sigma^2_q}\right)^{-1}, \qquad
\thetaq{a}{k} = \Sigmaq{a}{k} \left(\frac{\mu_{0,q}}{\sigma^2_{0,q}} + \frac{S^{a,k}_t}{\sigma^2_q}\right).
\end{equation}
The C-posterior uses the analogous form with $(\thetac{a}{k}, \Sigmac{a}{k})$, $\bar c_a$ in place of $\mu_{0,q}$, $\sigma^2_c$ in place of $\sigma^2_q$, and a separate pair of sufficient statistics over realized costs.

\paragraph{Why a learned cost posterior.}
In stationary settings with known prices, a fixed scalar $\bar c_a$ would suffice for cost. The agent-native regime drops both assumptions: prices can change mid-run (S3 in our benchmark), and the agent only sees prices through quoted observations from the market. Treating cost as a Gaussian posterior over realized $c_t$ lets the policy detect price shocks the same way the Q-posterior detects quality shocks---through the effective sample count $n^{a,k}_t$ and the discount that reweights stale observations. We demonstrate empirically (\S\ref{sec:eval-ablations}) that disabling this and pinning cost to $\bar c_a$ collapses adaptation to S3.

\subsection{Discounted sufficient statistics}
\label{sec:method-discount}

We apply a multiplicative discount $\gamma \in (0, 1]$ to every cell's $n$ and $S$ at the start of each round, before any new observation arrives:
\begin{equation}
n^{a,k}_t = \gamma \cdot n^{a,k}_{t-1}, \qquad
S^{a,k}_t = \gamma \cdot S^{a,k}_{t-1}.
\end{equation}
After the round's chosen arm produces an observation, only its bucket increments:
\begin{equation}
n^{a_t, k_t}_t \mathrel{+}= 1, \quad
S^{a_t, k_t}_t \mathrel{+}= \ut\,\,(\text{Q-posterior}), \quad
{S_c}^{a_t, k_t}_t \mathrel{+}= c_t\,\,(\text{C-posterior}).
\end{equation}
Discounting applies regardless of whether a cell was pulled this round---non-stationarity is wall-clock, not pull-conditional. We use $\gamma = 0.999$ for both posteriors (effective half-life $\approx 693$ rounds), chosen to react cleanly within the shock windows we evaluate (S2 covers $2{,}500$ rounds, S3 has $9{,}000$ post-shock rounds) without forgetting the stationary baseline too quickly. Sensitivity to $\gamma$ is in [/cc TODO sensitivity appendix].

\subsection{Decision rule}
\label{sec:method-decision}

At each round, the wallet exposes $\lambda_t$ and the affordable set $\mathcal{A}_t$. The policy reads $\lamn$ from the wallet, samples both posteriors per affordable arm, and ranks them:
\begin{equation}
\label{eq:padct-sample}
\hatu{a}{k_t} \sim \mathcal{N}\!\big(\thetaq{a}{k_t},\, \Sigmaq{a}{k_t}\big), \qquad
\hatc{a}{k_t} \sim \mathcal{N}\!\big(\thetac{a}{k_t},\, \Sigmac{a}{k_t}\big),
\end{equation}
\begin{equation}
\label{eq:padct-score}
\tilde{\hat c}^{a,k_t} = \mathrm{clip}\!\big(\hatc{a}{k_t}/c_{\max},\, 0,\, 1\big), \qquad
\pi_a = (1 - \lamn)\,\hatu{a}{k_t} - \lamn\,\tilde{\hat c}^{a,k_t}.
\end{equation}
The chosen arm is $a_t = \arg\max_{a \in \mathcal{A}_t} \pi_a$. Algorithm~\ref{alg:padct} fixes the per-round procedure end-to-end, with each block annotated by the PA-DCT component it implements.

\begin{algorithm}[t]
\caption{PA-DCT (Payment-Aware Discounted Contextual Thompson sampling). Block annotations P, D, C, TS mark the four ablatable components.}
\label{alg:padct}
\begin{algorithmic}[1]
\Require Provider set $\{1,\ldots,K\}$ with spec costs $\bar c_a$; discount $\gamma$; failure penalty $\nu$; cost normalizer $c_{\max}$; encoder $k(\cdot)$; priors $(\mu_{0,q}, \sigma^2_{0,q}, \sigma^2_q)$ and $(\bar c_a, \sigma^2_{0,c}, \sigma^2_c)$; wallet sharpness $\alpha$.
\For{$(a, k) \in \{1,\ldots,K\} \times \{1,\ldots,K_b\}$}
  \State init Q-posterior$_{a,k} \leftarrow \mathcal{N}(\mu_{0,q}, \sigma^2_{0,q})$;\quad C-posterior$_{a,k} \leftarrow \mathcal{N}(\bar c_a, \sigma^2_{0,c})$
\EndFor
\For{$t = 0, 1, \ldots, T-1$}
  \State \emph{(D)} discount every Q- and C-cell by $\gamma$
  \State observe $\mathbf{x}_t$, $B_t$;\quad $k_t \leftarrow k(\mathbf{x}_t)$
  \State \emph{(P)} $\lambda_t \leftarrow \exp(\alpha \cdot \bex_t)$;\quad $\lamn \leftarrow \lambda_t / (1 + \lambda_t)$
  \State $\mathcal{A}_t \leftarrow \{a : B_t \geq \mathrm{effective\_price}(a, t)\}$
  \If{$\mathcal{A}_t = \emptyset$} halt (bankruptcy) \EndIf
  \For{$a \in \mathcal{A}_t$}
    \State \emph{(TS, C)} sample $\hatu{a}{k_t} \sim$ Q-posterior$_{a, k_t}$;\quad $\hatc{a}{k_t} \sim$ C-posterior$_{a, k_t}$
    \State $\tilde{\hat c}^{a,k_t} \leftarrow \mathrm{clip}(\hatc{a}{k_t}/c_{\max}, 0, 1)$;\quad $\pi_a \leftarrow (1-\lamn)\,\hatu{a}{k_t} - \lamn\,\tilde{\hat c}^{a,k_t}$
  \EndFor
  \State $a_t \leftarrow \arg\max_{a \in \mathcal{A}_t} \pi_a$
  \State pay via the protocol; observe $(q_t, c_t, f_t)$;\quad $\ut \leftarrow q_t - \nu f_t$
  \State update Q-posterior$_{a_t, k_t}$ with $\ut$;\quad update C-posterior$_{a_t, k_t}$ with $c_t$
  \State wallet.\texttt{record\_spend}$(c_t)$
\EndFor
\end{algorithmic}
\end{algorithm}

\subsection{Hyperparameters}
\label{sec:method-hyperparams}

We fix the following constants across every experiment in the paper:
\begin{itemize}
\item Quality posterior: $\mu_{0,q} = 0.5$, $\sigma^2_{0,q} = 1.0$, $\sigma^2_q = 0.09$ (allowing substantial heterogeneity in observed quality).
\item Cost posterior: $\sigma^2_{0,c} = 10^{-4}$, $\sigma^2_c = 10^{-6}$ (tight, because realized cost is nearly deterministic given the provider and scenario).
\item Discounts: $\gamma_q = \gamma_c = 0.999$.
\item Reward weights and bandit constants: $\nu = 0.5$, $\alpha = 2$, $c_{\max} = \$0.01$.
\item Bucket count: $K_b = 4$ (one per task type, see \S\ref{sec:eval-setup}).
\end{itemize}
None of these are tuned per scenario or per seed.

\subsection{Theoretical properties}
\label{sec:method-theory}

PA-DCT is a bandit-family method. Its constituent components---contextual Thompson sampling~\citep{russo2014learning}, discounted sufficient statistics for non-stationary bandits~\citep{garivier2011discounted}, and budget-aware reward shaping in the spirit of bandits-with-knapsacks~\citep{badanidiyuru2018bandits}---are each individually well-studied. Our analysis confirms that the specific combination instantiated by PA-DCT, which to our knowledge has not been analyzed jointly before in the agent-economy regime, preserves the standard properties expected of its parts.

Specifically, in Appendix~\ref{appendix:proofs} we establish four properties: (i) the payment-aware reward is bounded, $\rt \in [-1, +1]$ (Prop.~\ref{prop:reward-bounded}); (ii) per-cell posteriors are almost-surely consistent in the stationary regime (Prop.~\ref{prop:consistency}); (iii) under a well-specified Gaussian Bayesian model and $\gamma=1$, PA-DCT inherits the standard contextual Gaussian-TS Bayesian regret bound of order $\tilde O(\sqrt{K K_b T})$ via reduction to \citet{russo2014learning} Theorem~1 (Thm.~\ref{thm:stationary-regret}); and (iv) after an abrupt change-point of magnitude $\Delta$, the per-arm Q-posterior bias admits an exact closed-form bound, yielding a recovery horizon $\tau \geq \log(\Delta/\delta)/\log(1/\gamma)$ for any target tolerance $\delta$ in the data-dominated regime (Thm.~\ref{thm:adaptation}, Cor.~\ref{cor:recovery}). The discount factor $\gamma$ thus controls an explicit per-arm responsiveness to drift; for $\gamma = 0.999$ the effective horizon is roughly $1{,}000$ rounds (Lemma~\ref{lem:ess}).

We do not establish a coupled regret bound for piecewise-stationary regimes with budget pressure, nor a misspecified-likelihood version of Thm.~\ref{thm:stationary-regret}; full theorem statements, proofs, and limitations of this analysis are deferred to Appendix~\ref{appendix:proofs} (limitations in App.~\ref{app:proof-limitations}).
