\section{Discussion and Limitations}
\label{sec:discuss}

\paragraph{Pre-payment vs post-selection.}
402Pilot is a pre-payment decision layer: it picks before the HTTP request fires. A complementary line of work, including A402's atomic service channels~\citep{a402}, addresses post-selection issues such as undelivered services, dispute resolution, and atomic refund. The two are layered, not in tension: 402Pilot tells the agent which provider to engage, and protocols below the layer guarantee that the engagement is settled atomically. We see no fundamental obstacle to running 402Pilot above an A402-style channel; pre-payment selection and post-payment atomicity address different failure modes.

\paragraph{Threats to validity.}
We are explicit about the gaps between 402Pilot-Bench and a live x402 deployment; reviewers and practitioners should weigh each before generalizing our results.

\emph{(a) Frozen replay.} Every result in this paper comes from a deterministic replay of $20{,}575$ pre-generated responses; no LLM was queried during the bandit experiments. This buys reproducibility and amortizes inference cost across $30$ seeds, but it freezes provider behaviour at the $V=5$ versions per (task, provider) we sampled and cannot exhibit drift faster than the replay buffer. Two mitigations are intrinsic to PA-DCT---the discount $\gamma$ downweights stale observations, the cost posterior tracks realized prices---but neither substitutes for re-pre-generation in long-running deployments. We do not claim our replay results extrapolate to live markets without re-validation; live evidence with a real x402 facilitator is the most important next step.

\emph{(b) LLM-as-judge on T3b.} The open-ended web-search bucket scores quality with a cached LLM-as-judge against a structured rubric. The judge accepts fluent-but-wrong P-adv answers more often than a deterministic evaluator does, so PA-DCT identifies P-adv on T3b more slowly and less completely than on T1/T2 (\S\ref{sec:eval-main}). This is an \emph{evaluator} boundary, not a \emph{policy} one; with a tighter judge, a verifier in the loop, or human-in-the-loop spot checks, the gap should close. Our $\bar q$ on T3b is therefore a judge-confounded upper bound on truthfulness.

\emph{(c) Non-strategic providers.} We assume providers post a price, deliver a service, and accept the buyer's outcome score. Strategic providers, considered by~\citet{aamas2026reverseauction}, are out of scope and would change the cost-posterior interpretation: realized cost would become a strategic action rather than a market signal. Combining a 402Pilot-style decision layer with such mechanisms is plausible (the decision layer would consume the auction's allocations as quoted prices) but lies outside this paper's scope.

\emph{(d) Single-wallet, single-buyer; no concurrent reserve.} PA-DCT addresses one wallet, one policy, one workload stream. We do not model concurrent agents drawing from a shared reserve, partial-fill provider behaviour under heavy load, rate-limit interaction between several 402Pilot instances on the same provider, or multi-agent learning effects. Cross-agent reserve contention can change the affordable set in ways our wallet abstraction does not capture.

\emph{(e) No live x402 settlement.} We do not measure end-to-end latency, transaction-fee jitter, or settlement failures from a real x402 facilitator; the loop assumes payments either succeed atomically or fail loudly. Live settlement could introduce additional failure modes (mempool delay, replacement attacks, facilitator outage) that \emph{interact} with budget pressure rather than being absorbed by it---e.g., a settlement-failure burst that drains $\lambda_t$ without delivering quality would push the reward calculation off the regime where Prop.~\ref{prop:reward-bounded} is calibrated.

\emph{(f) Provider quality ordering within task type.} Contextual bucketing assumes provider quality can vary across the four task-type buckets but does not stratify within type along finer dimensions. For our calibrated providers, model scale (parameter count) is the dominant differentiator, so this is adequate; heterogeneous skill profiles (e.g., a small specialized provider that beats GPT-5.4 on a narrow domain) would benefit from finer-grained context features or a learned encoder. We leave this as future work.

We list these threats explicitly so that the empirical claims in \S\ref{sec:eval} are read against the regime they were measured in.

\paragraph{What we did not study but flag.}
\emph{(i)} Drift detection: PA-DCT relies on $\gamma$ for non-stationarity but does not run an explicit change-point detector. Adaptive $\gamma$ tied to a CUSUM-style detector is a natural next step. \emph{(ii)} Trust signals: provider attestations from A402-style channels could feed the prior of the Q- or C-posterior; we did not explore this. \emph{(iii)} Cold start: in deployment with new providers, our priors fall back to the spec sheet; principled cold-start strategies (transfer, meta-learning) are not addressed here. \emph{(iv)} Fairness/safety constraints over provider sets (e.g., regional regulation, content policy): out of scope; the policy currently trusts the affordable set as a black box.
