\section{Conclusion}
\label{sec:conclusion}

We formalized \emph{agent-native payment decision-making}: budget-aware, non-stationary, contextual provider selection in programmable micropayment markets where the buyer is an autonomous wallet-bearing agent, not a human-operated routing service. The central claim that follows from this formalization is that autonomous payment is not only a settlement problem; it is an online economic decision problem in which the buyer must learn both service quality and price simultaneously. In markets where provider quality or price may drift, this pushes the buyer-side policy toward wallet-, context-, and shock-aware adaptation rather than a fixed provider rule. We proposed \emph{402Pilot}, a protocol-agnostic buyer-side decision layer that separates this decision from settlement and runs above x402 (or AP2, A2A--x402, A402, any future quote-and-pay protocol). Inside the layer we proposed \emph{PA-DCT}, a payment-aware Thompson-sampling rule built from four conceptual components (payment-awareness, discount, contextual bucketing, Thompson sampling), where payment-awareness is grounded in a \emph{learned cost posterior} that lets a single policy detect price shocks and quality shocks symmetrically.

On 402Pilot-Bench---a reproducible benchmark of $823$ tasks $\times$ $5$ providers (including same-price adversarial and flaky variants) $\times$ $3$ calibrated market scenarios with $30$ seeds and $10{,}000$ rounds per cell---PA-DCT recovers from a mid-tier outage $35\%$ faster than its no-discount variant, adapts within $\approx\!200$ rounds to a premium price promotion, and wins $13$ of $14$ non-stationary baseline comparisons on cumulative payment-aware reward; the remaining comparison is a statistically non-significant tie against Always-P-cheap in S2, where PA-DCT delivers materially higher mean quality. In the stationary scenario, PA-DCT incurs a measured exploration cost in exchange for adaptation capacity that the stationary market does not expose. The four-component ablation, plus a diagnostic isolation of the cost-posterior pathway, shows each ingredient is exposed on a different axis: payment-awareness on cumulative reward (uniformly necessary), discount on adaptation time, contextual bucketing selectively when shock structure is task-type heterogeneous, Thompson sampling on seed variance and shock-detection reliability, and the cost posterior on the price-shock channel of S3. No single classical metric reveals all of these; this motivates the five-metric framework (cum-PA, ROI, $\bar q$, regret, AdaptT) we recommend for the regime.

Code, the frozen pre-generated dataset, the scenario sweeps, and the evaluation harness are released as the artifact accompanying this paper (URL withheld during anonymous review; an anonymized snapshot is included in the supplemental material). We hope the layer interfaces, the algorithm, and the benchmark together invite extension to other agent-native payment decision-making settings beyond the LLM-buyer instance studied here.
