\section{Hyperparameter Sensitivity}
\label{appendix:sensitivity}

PA-DCT exposes seven scalar hyperparameters, all locked across every
experiment in the paper (\S\ref{sec:method-hyperparams}, \S\ref{sec:problem}).
This appendix collects what we know about robustness to each, separates
calibration-derived values from convention-derived values, and is honest
about which sweeps were run and which were not.

\subsection{Targeted S1 sensitivity sweep}
\label{app:s1-sensitivity}

To check that the locked defaults are calibrated rather than tuned, we run
a one-at-a-time perturbation of each scalar that PA-DCT exposes
($\gamma, \nu, \alpha, c_{\max}$), $10$ seeds per cell at $T = 10{,}000$
on S1, with a matching True Oracle in every cell so $\textsc{PA-gap}/T$
remains comparable across perturbations (the oracle's objective uses the
same $\nu$ and $c_{\max}$ as the PA-DCT cell under test).

\begin{table}[H]
\centering
\scriptsize
\setlength{\tabcolsep}{2.2pt}
\renewcommand{\arraystretch}{0.98}
\caption{One-at-a-time S1 sensitivity of PA-DCT. Means $\pm$ SE over $10$
seeds; each non-default row perturbs a single scalar, the rest stay at
their default values.}
\label{tab:s1-sensitivity}
\begin{tabular}{@{}lrrrr@{}}
\toprule
Setting & $\bar q_T$ & Budget \% & ROI & $\textsc{PA-gap}/T$ \\
\midrule
\textsc{default}       & $0.796{\pm}0.001$ & $42.2{\pm}0.3$ & $377.5{\pm}2.1$ & $+0.1328{\pm}0.0013$ \\
\midrule
$\gamma=0.99$      & $0.753{\pm}0.001$ & $49.8{\pm}0.2$ & $302.4{\pm}1.3$ & $+0.2087{\pm}0.0013$ \\
$\gamma=0.9995$    & $0.801{\pm}0.002$ & $40.8{\pm}0.4$ & $392.6{\pm}3.3$ & $+0.1218{\pm}0.0016$ \\
$\nu=0.25$         & $0.795{\pm}0.001$ & $42.1{\pm}0.4$ & $377.6{\pm}3.0$ & $+0.1329{\pm}0.0018$ \\
$\nu=1.0$          & $0.797{\pm}0.001$ & $42.1{\pm}0.3$ & $378.7{\pm}2.3$ & $+0.1324{\pm}0.0014$ \\
$\alpha=1.0$       & $0.788{\pm}0.002$ & $36.7{\pm}0.3$ & $430.3{\pm}3.1$ & $+0.1045{\pm}0.0011$ \\
$\alpha=4.0$       & $0.802{\pm}0.001$ & $54.7{\pm}0.4$ & $293.5{\pm}2.0$ & $+0.1781{\pm}0.0016$ \\
$c_{\max}=0.005$   & $0.787{\pm}0.001$ & $41.9{\pm}0.4$ & $375.9{\pm}3.0$ & $+0.1567{\pm}0.0028$ \\
$c_{\max}=0.02$    & $0.800{\pm}0.001$ & $50.5{\pm}0.3$ & $316.8{\pm}1.9$ & $+0.1490{\pm}0.0021$ \\
\bottomrule
\end{tabular}
\end{table}

No cell bankrupts the wallet, and $\bar q_T$ stays inside the narrow
band $[0.753, 0.802]$; the relative ordering $\textsc{PA-gap}/T > 0$
holds in every cell. The two perturbations that matter are the two that
the algorithm exposes as scheduling levers: $\gamma$ (memory horizon)
and $\alpha$ (budget-pressure sharpness); $\nu$ and $c_{\max}$ move
$\textsc{PA-gap}/T$ by less than one third of $\gamma$'s effect within
the tested range. The $c_{\max}$ cells also confirm that fixing
$c_{\max}=\$0.01$ to the highest spec price in Table~\ref{tab:providers}
is not decisive: halving or doubling it moves Budget\% by $\pm 8$pp but
leaves $\bar q_T$ within $1.5$pp of default. The
$\gamma_{\mathrm{low}}{=}0.99$ cell doubles
$\textsc{PA-gap}/T$ on S1, which is the ``cost'' half of the
short-memory--fast-recovery trade-off whose ``recovery'' half appears in
Table~\ref{tab:gamma-sensitivity} and the $-D$ ablation in
Table~\ref{tab:ablations}.

\subsection{\texorpdfstring{Discount factor $\gamma$}{Discount factor gamma}}
\label{app:sens-gamma}

Table~\ref{tab:s1-sensitivity} gives the targeted S1 sweep around the
locked default. For cross-scenario evidence, we also compare full PA-DCT
($\gamma = 0.999$) against the $-D$ ablation ($\gamma = 1$) across all
three scenarios (Table~\ref{tab:gamma-sensitivity}, drawn from
Table~\ref{tab:ablations}).

\begin{table*}[t]
\centering
\small
\caption{Two-point sensitivity to the discount factor $\gamma$ from the
$-D$ ablation. Means over $30$ seeds. The pattern
across scenarios is consistent with the analytical view in
Cor.~\ref{cor:recovery}: $\gamma < 1$ pays a small cost in stationary
settings and earns it back via faster recovery under shocks.}
\label{tab:gamma-sensitivity}
\resizebox{\linewidth}{!}{%
\begin{tabular}{l|cccc|cccc|cccc}
\toprule
& \multicolumn{4}{c|}{S1 Stationary} & \multicolumn{4}{c|}{S2 Mid Outage} & \multicolumn{4}{c}{S3 Premium Promo} \\
$\gamma$ & PA-gap/$T$ & ROI & $\bar q_T$ & AdaptT & PA-gap/$T$ & ROI & $\bar q_T$ & AdaptT & PA-gap/$T$ & ROI & $\bar q_T$ & AdaptT \\
\midrule
$0.999$ & $0.133$ & $378$ & $0.797$ & \texttt{n/a} & $0.166$ & $357$ & $0.761$ & $1467$ & $0.121$ & $429$ & $0.831$ & $200$ \\
$1.000$ & $0.109$ & $412$ & $0.810$ & \texttt{n/a} & $0.170$ & $399$ & $0.745$ & $2249$ & $0.114$ & $426$ & $0.838$ & $279$ \\
\bottomrule
\end{tabular}%
}
\end{table*}

The cross-scenario endpoints sandwich the regimes that matter for our analysis:
$\gamma = 1$ is the no-discount fallback that recovers a vanilla contextual
TS, and $\gamma = 0.999$ is the value chosen so the effective horizon
$1/(1-\gamma) = 10^3$ rounds is short enough to react inside the S2 outage
window ($2{,}500$ rounds) yet long enough to retain useful evidence in S1.
The S1 sweep adds a local check at $\gamma=0.99$ and $\gamma=0.9995$:
short memory sharply worsens stationary PA-gap/$T$, while longer memory
improves S1 slightly. Adaptive-$\gamma$ strategies are flagged as future
work in \S\ref{sec:discuss}. Cor.~\ref{cor:recovery} provides the
analytical counterpart: in the data-dominated regime, the bias decays as
$\Delta\,\gamma^{\tau}$, so smaller $\gamma$ buys faster recovery at fixed
$\tau$ at the cost of higher steady-state variance --- a tradeoff visible
in the table above (S2 AdaptT improves $35\%$ at the cost of a larger S1
PA-gap/$T$).

\subsection{\texorpdfstring{Failure penalty $\nu$}{Failure penalty nu}}
\label{app:sens-nu}

We use $\nu = 0.5$, chosen to abstract the structural cost of timeouts ---
retry overhead, broken downstream call chains, and agent stalling --- beyond
the trivial fact that the quality score $q_t$ is already $0$ on a parse
failure. The role of this penalty is concentrated entirely on P-flaky,
whose only difference from P-mid is its timeout rate; without a non-zero
$\nu$, P-flaky is indistinguishable from P-mid on the failure-free subset
of rounds, and the policy has no signal that would let it migrate spend
away once flakiness is observed. Table~\ref{tab:s1-sensitivity} (the
$\nu=0.25$ and $\nu=1.0$ cells) confirms that
$\nu \in \{0.25, 1.0\}$ moves $\bar q_T$ and $\textsc{PA-gap}/T$ by under
$0.5\%$ on S1, where P-flaky failures are infrequent and bounded. A
natural direction for follow-up work is to learn $\nu$ per provider from
observed downstream stalls rather than treat it as a global constant.

\subsection{\texorpdfstring{Budget-pressure sharpness $\alpha$}{Budget-pressure sharpness alpha}}
\label{app:sens-alpha}

The budget-pressure schedule
$\lambda_t = \lambda_0\exp(\alpha\cdot \bdev_t)$ is parameterized by $\alpha = 2$ with $\lambda_0=1$.
With $\alpha$ smaller, $\lambda_t$ rises slowly even when the agent is
spending well over plan, and the cost penalty in $\rt$ kicks in too late
to alter behavior; with $\alpha$ larger, $\lambda_t$ saturates the sigmoid
weight $\lamn$ at the first hint of overshoot and the policy collapses
onto the cheapest arm regardless of utility. The chosen value is the
midpoint of an order-of-magnitude calibration band, and is paired with
$\mathrm{target\_rate} = 1/T = 10^{-4}$ (the spend fraction per round
that exhausts $B$ exactly at $T$); this calibration locks $\lamn = 0.5$
when on plan and pushes $\lamn \to 1$ smoothly under sustained
overshoot. Table~\ref{tab:s1-sensitivity} quantifies the two-sided
trade-off: $\alpha{=}1$ underspends the wallet (Budget\% drops from $42$
to $37$) and yields a smaller PA-gap, while $\alpha{=}4$ pays $13$pp
more budget for a $34\%$ larger PA-gap. The $-P$ ablation
($\lamn \equiv 0$, equivalent to $\alpha \to 0$) is the one-sided
endpoint that disables budget pressure entirely; see Table~\ref{tab:ablations}.

\subsection{Posterior priors and likelihood variances}
\label{app:sens-priors}

The Q-posterior uses prior $\mathcal{N}(\mu_{0,q}, \sigma^2_{0,q}) =
\mathcal{N}(0.5, 1.0)$ and likelihood variance $\sigma^2_q = 0.09$; the
C-posterior uses prior $\mathcal{N}(\bar c_a, \sigma^2_{0,c}) =
\mathcal{N}(\bar c_a, 10^{-4})$ and likelihood $\sigma^2_c = 10^{-6}$.
The Q-posterior values are the standard ``broad mid-range prior, moderate
proxy noise'' choice for a $[0,1]$-bounded reward (cf.\
\citealp{lattimore2020bandit}, Ch.~36); in our setting they are wide
enough that the Bayesian model is mildly conservative (cf.\ misspecification
discussion in App.~\ref{app:proof-limitations}) and the policy explores
adequately in the early rounds of every scenario. The C-posterior values
reflect the empirical structure of the cost channel: realized $c_t$ is
nearly deterministic given the provider and the round (only the S3
premium-drop transition, the only price shock in our benchmark, makes it
non-degenerate), so a tight prior plus tighter likelihood lets the
posterior follow the spec sheet at calibration time and migrate quickly
once the shock arrives. We did not sweep these four scalars; the
$-TS$ ablation rules out the qualitative role of variance entirely
(replacing samples by means; see Table~\ref{tab:ablations} for the
$5$--$9\times$ seed-variance penalty), so the residual question is one
of quantitative tuning rather than structural design.

\subsection{Constants we explicitly do not sweep, and why}
\label{app:sens-not-swept}

We deliberately do not sweep the locked constants of \S\ref{sec:problem}
($T$, $B$). $T = 10{,}000$ is the run length built into every
seeded trajectory in the released artifact; running shorter is a strict
sub-experiment of running longer. $B = \$50$ is calibrated against the
strongest non-adaptive baseline (Always-P-premium exhausts at $\approx
5{,}000$ rounds), so the budget cliff falls in the second half of every
run --- changing $B$ would shift this cliff and re-calibrate every
baseline along with PA-DCT.
A sweep over $(T, B)$ that preserves the budget-cliff timing is mechanically
equivalent to scaling time, and hence does not generate new information
about PA-DCT itself.
