\section{402Pilot-Bench Dataset Card}
\label{appendix:dataset}

402Pilot-Bench is the replay benchmark introduced as Contribution C3
(\S\ref{sec:intro}); it is the artifact every empirical claim in this
paper is computed against. This appendix is a datasheet
in the spirit of \citet{gebru2021datasheets}, covering composition,
collection, evaluation, intended use, and limitations. We treat the
benchmark as a first-class deliverable: the release ships with
deterministic seed and split protocols so that other policies can be
dropped into the same harness and produce numbers comparable to ours.

\subsection{Composition}
\label{app:dataset-composition}

\paragraph{Tasks.}
402Pilot-Bench contains $823$ unique tasks drawn from three public
sources plus one custom subset, partitioned into four evaluation
buckets used by PA-DCT's contextual structure ($K_b = 4$). Counts and
sources are reported in Table~\ref{tab:dataset-card}.

\begin{table*}[t]
\centering
\small
\caption{Per-bucket composition of 402Pilot-Bench. The web-search domain
splits into closed-form (T3a) and open-ended (T3b) sub-buckets because
the two use different evaluators (deterministic max EM/F1 vs.\ cached
LLM-as-judge). One TriviaQA item is excluded across all providers due
to a deterministic content-filter rejection on the cheapest provider's
API; excluding it for every provider keeps quality comparisons
apples-to-apples.}
\label{tab:dataset-card}
\begin{tabularx}{\linewidth}{@{}llXrl@{}}
\toprule
Bucket & Domain & Source & Tasks & License \\
\midrule
T1  & Coding             & HumanEval~\citep{chen2021humaneval}             & $164$ & MIT \\
T2  & Multi-hop QA       & HotpotQA~\citep{yang2018hotpotqa}               & $220$ & CC BY-SA 4.0 \\
T3a & Web search (closed)& TriviaQA web split~\citep{joshi2017triviaqa}    & $219$ & Apache-2.0 \\
T3b & Web search (open)  & Custom open-ended subset (this work)            & $220$ & Same as artifact \\
\midrule
\multicolumn{3}{l}{\textbf{Total}} & $\mathbf{823}$ & \\
\bottomrule
\end{tabularx}
\end{table*}

\paragraph{Pre-generated responses.}
For each (task, provider) pair we pre-generate $V = 5$ response versions,
yielding $823 \times 5 \times 5 = 20{,}575$ total responses. Each response
is stored on disk with the schema \texttt{(schema\_version, task\_id,
task\_type, provider\_id, version, response, cost\_usdc, latency\_s,
failure\_flag, failure\_code, quality\_score, generated\_at, temperature)}
in JSONL, with one file per (provider, bucket) pair. Quality scores are
computed once at pre-generation time and cached, so all bandit results
are deterministic given the seed.

\paragraph{Sampling.}
HumanEval includes all $164$ canonical tasks. The HotpotQA, TriviaQA, and
open-ended subsets are sampled from their source datasets using
NumPy-default RNG with a fixed seed (constant in the loader source);
sampled indices are sorted and recorded in the per-source cache file so
the sample is reproducible from the cache regardless of upstream
dataset changes. The TriviaQA exclusion is a single-task content-filter
skip-list applied at load time across all providers; the on-disk cache
retains the full $220$-task sample, and removing the entry from the
skip list restores it without rebuilding.

\subsection{Provider configuration}
\label{app:dataset-providers}

The five providers are summarized in Table~\ref{tab:providers} of the
main text and instantiated in \texttt{experiments/main.yaml}. We
emphasize three design points:

\begin{itemize}
\item P-mid, P-adv, and P-flaky share the same base price (\$0.002/call)
  and base model. They are distinguishable only by behavior --- correct
  answers (P-mid), fluent-but-wrong answers (P-adv), and a $40\%$
  timeout-injection rate (P-flaky) --- and not by any static metadata. A
  policy that ranks providers by spec sheet alone cannot tell them apart.
\item P-cheap (\$0.0005/call, weakest model) and P-premium (\$0.01/call,
  strongest model) anchor the cost--quality frontier and
  participate in S3, where P-premium's effective price drops to \$0.002
  at round $1{,}000$ to create the promotional opportunity.
\item Adversarial behavior (P-adv) is implemented through a system-prompt
  modification, not by changing the underlying model; flaky behavior
  (P-flaky) is implemented through a probabilistic timeout wrapper. Both
  are deterministic given the per-call seed.
\end{itemize}

\subsection{Evaluation protocol}
\label{app:dataset-eval}

Each bucket carries its own evaluator, fixed at pre-generation time.

\begin{itemize}
\item \textbf{T1 Coding (HumanEval).} pass@1 over the canonical test
  cases. Deterministic; quality $q_t \in \{0, 1\}$.
\item \textbf{T2 Multi-hop QA (HotpotQA).} Maximum of EM and F1 against
  the gold answer string. Deterministic given the gold answer; quality
  $q_t \in [0, 1]$.
\item \textbf{T3a Web search closed-form (TriviaQA-web).} Maximum of EM
  and F1 against the canonical answer plus the alias set provided by
  TriviaQA. Deterministic; quality $q_t \in [0, 1]$.
\item \textbf{T3b Web search open-ended.} LLM-as-judge with a fixed
  rubric (criteria: factual correctness, completeness, conciseness),
  prompted against a strong reference judge model. The judge is queried
  once at pre-generation per (task, provider, version) tuple and the
  resulting quality score is cached on disk; bandit replays read from
  the cache and never re-query the judge. Cached quality scores
  $q_t \in [0, 1]$ are persisted in a separate \texttt{judge\_cache.jsonl}
  file alongside the per-provider response files.
\end{itemize}

\paragraph{Failure flag.} Independently of the quality score, every
response carries a binary \texttt{failure\_flag}. The flag is set when
the provider call times out, the response cannot be parsed, or the
upstream API rejects the request. For P-flaky the flag fires
probabilistically by design; for the other providers it surfaces only
genuine API-side failures. Utility is $\ut = q_t - \nu f_t$ with
$\nu = 0.5$ (\S\ref{sec:problem}); the bandit treats failures as a
distinct signal from low quality.

\subsection{Intended use}
\label{app:dataset-intended-use}

402Pilot-Bench is intended for benchmarking buyer-side decision policies
in the agent-native payment decision-making setting: budget-aware, non-stationary,
contextual provider selection over a small set of paid LLM endpoints.
The release decouples bandit evaluation from inference noise (responses
are frozen) and lets a new policy be compared against ours on identical
seed paths in a few CPU-hours. Out of scope: protocol-level questions
(settlement, dispute resolution), strategic-provider mechanism design
(\S\ref{sec:discuss}), and any setting that requires live LLM calls
during the bandit loop. We do not recommend this benchmark for
training-time decisions or for studies that need within-trial
counterfactual access to non-chosen arms; both are outside the
single-call replay model.

\subsection{Limitations and known issues}
\label{app:dataset-limitations}

\paragraph{Frozen responses.} Pre-generation captures provider behavior
at calibration time. Real LLM markets evolve on timescales shorter than
a research cycle; live deployment should be preceded by re-pre-generation
against the deployment-time model versions. The discount $\gamma$ helps
the policy adapt within a fixed dataset, but it is not a substitute for
refreshing the dataset.

\paragraph{Evaluator-bounded P-adv detection on T3b.} The LLM-as-judge
on the open-ended sub-bucket occasionally accepts fluent-but-wrong
answers from P-adv. PA-DCT therefore identifies P-adv on T3b more slowly
and less completely than on the deterministic-evaluator buckets
(\S\ref{sec:eval-main}). This is an evaluator limit, not a policy
limit; with a tighter judge, a verifier in the loop, or human spot
checks the gap should close.

\paragraph{Single content-filter exclusion.} The DashScope API used by
P-cheap rejects one TriviaQA question outright; we exclude that
question from \emph{every} provider's task set so the comparison stays
apples-to-apples. The total task count after exclusion is $823$ (i.e.\
$220 \to 219$ in T3a). The skip list and the rationale for each entry
are documented in the loader source.

\paragraph{Provider heterogeneity is calibrated, not real.} The five
providers are designed to expose specific phenomena (same-tier
quality differences, adversarial fluency, flakiness, price shocks) in a
controlled and reproducible way. Real-world provider mixes will differ;
we do not claim the calibrated set spans the deployment-time provider
distribution.

\paragraph{Single-language, single-modality.} All tasks are
English-language text-in / text-out. Multi-lingual, code-execution
sandboxed, and multi-modal extensions are natural follow-up benchmarks.

\subsection{Reproducing tables}
\label{app:dataset-reproducing-tables}

The released artifact separates replay generation from table
aggregation. Recomputing tables from existing JSONL logs requires no live
LLM calls and usually takes minutes on a laptop; rerunning the full
$30$-seed, $10{,}000$-round sweeps can take several CPU-hours and is
dominated by local JSONL I/O.

Run the following scripts with \texttt{python -m scripts.<name>}:
\begin{itemize}
\item \texttt{compute\_main\_table} recomputes Table~\ref{tab:main}.
\item \texttt{compute\_significance} recomputes the paired checks in
  \S\ref{sec:eval-main}.
\item \texttt{compute\_ablation\_metrics} recomputes
  Table~\ref{tab:ablations}, including the S3 cost-posterior diagnostic.
\end{itemize}

To regenerate the JSONL logs before aggregation, use:
\begin{itemize}
\item \texttt{run\_scenario\_sweep} for the main table;
\item \texttt{run\_ablation\_matrix} for the four core ablation toggles;
\item \texttt{run\_s3\_promo\_ablation} for the S3 cost-posterior
  diagnostic, with the \texttt{padct} policy and existing S3 oracle logs.
\end{itemize}

\subsection{Distribution and maintenance}
\label{app:dataset-distribution}

The pre-generated responses, the loader code, the scenario classes, the
configuration YAML, and the scripts that regenerate every table and
figure in the paper are released as the supplemental artifact. Bug fixes
that change the on-disk cache will bump a \texttt{schema\_version} field;
fixes that change only the load-time filter (such as the content-filter
skip list) leave the cache file untouched. The artifact is anonymized for
review; the post-acceptance repository will track issues and pull
requests for new tasks, providers, and scenarios.
