\section{Introduction}
\label{sec:intro}

A new family of protocols---x402~\citep{coinbase2025x402v2}, AP2~\citep{google2025ap2}, A2A--x402~\citep{a2ax402}, A402~\citep{a402}---now lets software agents pay for online services at machine speed. An agent holding a stablecoin wallet can discover paid endpoints, request a quote, settle on-chain, and receive a response in a single HTTP round trip. Per-call micropayments between autonomous agents and arbitrary services are no longer hypothetical; they are an emerging substrate for machine-to-machine commerce. But protocol-level support for \emph{executing} payments answers only half of the buyer's problem. The other half---\emph{which} provider to pay, given a finite wallet, uncertain service quality, and prices and reliability that drift over time---the protocol stack does not answer. This paper proposes \emph{402Pilot}, a decision layer that sits above any such protocol and supplies the missing buyer-side policy, and \emph{PA-DCT} (Payment-Aware Discounted Contextual Thompson sampling), the algorithm that drives it.

Consider an agent with a fixed budget $B$ that must serve a stream of paid tasks against a set of $K$ candidate providers. Quality and cost differ across providers, but not in transparent ways. The cheapest endpoint may be reliable for simple lookups but fail on multi-hop reasoning. The most expensive may be over-provisioned for trivial requests yet necessary for difficult ones. Providers at the same advertised price may differ silently in unspecified ways: one returns correct answers, another sounds fluent but is wrong, a third times out frequently. The x402 V2 specification recognizes this gap explicitly and leaves provider selection to the buyer ``based on historical outcomes''~\citep{coinbase2025x402v2}; its Discovery extension surfaces metadata that makes the choice possible but never specifies the choice itself. The agent must therefore learn, online, with only one observed outcome per paid call, how to allocate a wallet that drains. We argue that this is a learning problem in its own right, distinct from both classical bandit theory and recent LLM routing.

Three bodies of prior work touch our setting but do not cover it. Bandit theory provides the right ingredients individually: Budgeted Thompson Sampling for finite-spend constraints~\citep{xia2015bts}, Discounted Thompson Sampling for non-stationary rewards~\citep{qi2023dsts}, Contextual Thompson Sampling for task-conditional value~\citep{agrawal2013contextualts}, and bandits with knapsacks for the broader constrained-resource frame~\citep{badanidiyuru2018bandits}. Each, however, is studied in a setting where cost is a known scalar, the operator is a human researcher, and the experiment terminates cleanly. LLM cost-aware routing~\citep{chen2023frugalgpt,ong2024routellm,ding2024hybridllm,mixllm2024,madaan2024automix,llmbandit2025} addresses quality--cost tradeoffs over multiple models, but evaluation is offline against static rate cards, no payment protocol enters the decision loop, and prices do not change mid-run. Concurrent mechanism-design work on truthful reverse auctions for adaptive selection~\citep{aamas2026reverseauction} applies contextual bandits to provider choice but in a market where strategic providers submit bids and the designer optimizes incentive compatibility, not buyer adaptation. None of these settings is the one a single autonomous buyer actually faces against a quote-and-pay payment protocol.

We name this setting \emph{agent-native bandits}, distinguishing it from \emph{agent-implemented but human-operated} bandits, which describes most prior bandit work applied to agents. Three properties together define the regime. First, the operator of the bandit is the autonomous agent itself: there is no human team that resets state between market events, retunes hyperparameters after a price shock, or removes a degraded provider from the pool. The policy must remain coherent across the entire run. Second, the agent owns and watches a wallet that drains; budget pressure is a state observed and acted on at every round, not a static cap that triggers shutdown. Third, costs are first-class observations: the agent reads a quoted price, pays it, and must learn to predict prices the same way it learns to predict quality, because both can change. Naming this regime is not labeling for its own sake; each property forces a concrete design choice we make and ablate.

402Pilot is a protocol-agnostic decision layer with two stable interfaces: \texttt{select(task, affordable\_providers, wallet\_state)}, which returns a chosen provider, and \texttt{update(task, provider, observed\_cost, observed\_quality, failure)}, which feeds the realized outcome back. We instantiate the layer above x402 in this paper, but the same interface applies to AP2, A2A--x402, or any future quote-and-pay protocol; the layer never crosses the settlement boundary. Inside the layer, our policy PA-DCT carries four ingredients into the agent-native regime. (1) \textbf{Payment-aware} ranking: at decision time, sampled quality is offset by sampled normalized cost weighted by a sigmoid of wallet pressure $\lamn = \lambda_t/(1+\lambda_t)$, keeping the reward bounded in $[-1,+1]$ and tying decisions to wallet state without contaminating beliefs. (2) \textbf{Discounted} sufficient statistics: posteriors forget at rate $\gamma$, so observations from before a market shift weigh less than recent ones. (3) \textbf{Contextual} bucketing: a separate posterior per (provider, task-type) cell lets the agent learn that a provider is strong on coding but weak on multi-hop QA. (4) \textbf{Thompson sampling} of \emph{both} a quality posterior and a cost posterior per cell, so a price drop is detected the same way a quality drop is. The dual posterior over cost is the structural commitment that distinguishes PA-DCT from prior payment-aware bandit work and follows directly from agent-native property (3) above.

Throughout the paper we instantiate the regime with five providers, three task types, and three market scenarios. The five providers include three at the same nominal price tier of \$0.002/call: P-mid (capable, reliable), P-adv (a same-tier adversary that returns fluent but wrong answers), and P-flaky (a same-tier provider that times out 40\% of the time). They share base price and base model and therefore cannot be distinguished by any static rule; only reward feedback separates them. The three scenarios add controlled non-stationarity: \emph{S1} keeps the market stationary; \emph{S2} introduces a mid-tier outage from round 3{,}000 to round 5{,}500 in which P-mid fails 30\% of calls; \emph{S3} drops P-premium's price from \$0.01 to \$0.002 (matching the mid tier) at round 1{,}000, creating a promotional opportunity that a static policy cannot exploit and a budget-blind learner cannot afford. Each scenario probes a different facet of the regime: identity (S1), reliability shocks (S2), price shocks (S3).

We make five contributions, locked across the rest of the paper.

\textbf{C1. Setting characterization.} We formalize agent-native bandits and show, through a four-thread literature survey (\S\ref{sec:related-work}), that no existing line of work---agent payment protocols, classical bandits, LLM routing, or mechanism design over LLM providers---operates in this regime.

\textbf{C2. The PA-DCT algorithm.} We assemble a payment-aware Thompson-sampling rule for the agent-native regime. Quality and cost are tracked as separate Gaussian posteriors per (provider, task-type) bucket, both discounted; decision-time budget pressure enters through a bounded sigmoid weighting. Treating cost as a learned posterior rather than a known scalar is the structural commitment that lets one policy detect price shocks and quality shocks symmetrically (\S\ref{sec:method}).

\textbf{C3. The 402Pilot decision layer.} We articulate a protocol-agnostic interface between agent decision-making and payment execution. The layer commits to a clean separation: the protocol stack handles \emph{how} to pay; 402Pilot handles \emph{which} provider to pay (\S\ref{sec:layer}).

\textbf{C4. 402Pilot-Bench.} We construct a reproducible benchmark: 824 tasks across coding (HumanEval), multi-hop QA (HotpotQA), and web search (TriviaQA, custom open-ended); five providers including adversarial and flaky variants at the same price tier; three calibrated market scenarios (stationary, mid-tier outage, premium promotion); 30 seeds and 10{,}000 rounds per cell, totaling 20{,}575 pre-generated, frozen LLM responses (\S\ref{sec:eval-setup}). The benchmark decouples bandit evaluation from inference noise and is released as the artifact accompanying this paper.

<<<<<<< HEAD
\textbf{C5. Empirical findings.} On 402Pilot-Bench, PA-DCT outperforms five baselines on cumulative payment-aware reward under both non-stationary scenarios (S2 outage and S3 promo) with paired-bootstrap significance at $p<10^{-3}$ (\S\ref{sec:eval-main}). In the stationary scenario S1 it pays a small exploration cost relative to the strongest fixed baseline, offset by higher ROI on cheaper-task picks (\S\ref{sec:eval-main}). A four-component ablation (\S\ref{sec:eval-ablations}) shows that each of Payment-aware, Discount, Contextual, and Thompson sampling earns its place on a different evaluation axis; no single classical metric reveals all four, motivating the multi-metric framework we recommend.
=======
\textbf{C5. Empirical findings on which classical components transfer.} On 402Pilot-Bench, PA-DCT outperforms five baselines (Random, Always-Cheap, Always-Mid, Always-Premium, BudgetRule) on cumulative payment-aware reward in both non-stationary scenarios (S2 outage and S3 promo) with paired-bootstrap significance at $p<10^{-3}$ (\S\ref{sec:eval-main}); in the stationary baseline S1 it pays a small exploration cost relative to Always-Mid that a multi-metric view (\S\ref{sec:eval-main}) puts in context. A four-component ablation isolates each ingredient's empirical role (\S\ref{sec:eval-ablations}). Payment-awareness is uniformly necessary: without it cumulative reward collapses to negative across all three scenarios. Discounting is invisible on cumulative reward but speeds shock recovery by 35\% in S2 outage. Contextual bucketing helps when task-type heterogeneity is exploitable in the shock structure (S3) but not when shocks are uniform across task types (S2). Thompson sampling reduces seed variance by $5$--$9\times$ relative to greedy and enables shock detection that the greedy variant misses by an order of magnitude. No single classical bandit metric---not regret alone, not ROI alone---reveals all four contributions; this motivates the multi-metric evaluation framework we report.
>>>>>>> f4152aae22200d442744ab84c2288b439a163e6c

\S\ref{sec:related-work} positions 402Pilot against four threads of prior work. \S\ref{sec:problem} formalizes the round protocol, reward, and budget dynamics. \S\ref{sec:layer} describes the decision-layer architecture and its interfaces. \S\ref{sec:method} gives PA-DCT in detail with pseudocode. \S\ref{sec:eval} reports the main results, ablations, and limitations.
