\section{Discussion and Limitations}
\label{sec:discuss}

\paragraph{Pre-payment vs post-selection.}
402Pilot is a pre-payment decision layer: it picks before the HTTP request fires. A complementary line of work, including A402's atomic service channels~\citep{a402}, addresses post-selection issues such as undelivered services, dispute resolution, and atomic refund. The two are layered, not in tension: 402Pilot tells the agent which provider to engage, and protocols below the layer guarantee that the engagement is settled atomically. We see no fundamental obstacle to running 402Pilot above an A402-style channel; pre-payment selection and post-payment atomicity address different failure modes.

\paragraph{Dataset fidelity.}
Pre-generated responses freeze model behaviour at calibration time. Real LLM markets evolve: provider models update, prices change unannounced, and quality distributions drift on timescales shorter than a research cycle. Two mitigations are intrinsic to PA-DCT—the discount factor $\gamma$ down-weights stale observations, and the cost posterior tracks realized prices rather than spec sheets—but neither replaces re-pre-generation in long-running deployments. We do not claim our replay results extrapolate to live markets without re-validation; the most important next step in this line is live deployment evidence with a real x402 facilitator.

\paragraph{Evaluator-bounded P-adv detection.}
On open-ended T3b tasks, the LLM-as-judge occasionally accepts fluent-but-wrong answers from P-adv. PA-DCT therefore identifies P-adv on T3b more slowly and less completely than on T1/T2 (deterministic evaluators). This is an \emph{evaluator} limit, not a \emph{policy} limit; with a tighter judge, a verifier in the loop, or human-in-the-loop spot checks, the gap should close. We document this asymmetry in \S\ref{sec:eval-main} for honesty rather than treating it as a known result.

\paragraph{Single-agent scope.}
PA-DCT addresses one wallet, one policy, one workload stream. Multi-agent dynamics—cross-agent learning, federated posteriors, market-clearing under simultaneous bandit choices, externalities from concurrent buyers on the same provider—are real and out of scope. The mechanism-design line~\citep{aamas2026reverseauction} is one orthogonal extension; cross-agent posterior sharing under privacy constraints is another.

\paragraph{Provider quality ordering across task types.}
The contextual bucketing assumes provider quality can vary across the four task-type buckets but does not stratify within type along finer dimensions. For our calibrated providers, where model scale (parameter count) is the dominant differentiator, this is adequate. Heterogeneous skill profiles—e.g., a small specialized provider that beats GPT-5.4 on a narrow domain—would benefit from finer-grained context features or a learned context encoder; we leave this as future work.

\paragraph{Strategic providers.}
Our setting assumes non-strategic providers: they post a price, deliver a service, and accept the buyer's outcome score. The mechanism-design literature on truthful auctions~\citep{aamas2026reverseauction} addresses the strategic-provider regime. Combining a 402Pilot-style decision layer with such mechanisms is plausible (the decision layer would consume the auction's allocations as quoted prices) but is outside this paper's scope.

\paragraph{What we did not study but flag.}
\emph{(i)} Drift detection: PA-DCT relies on $\gamma$ for non-stationarity but does not run an explicit change-point detector. Adaptive $\gamma$ tied to a CUSUM-style detector is a natural next step. \emph{(ii)} Trust signals: provider attestations from A402-style channels could feed the prior of the Q- or C-posterior; we did not explore this. \emph{(iii)} Cold start: in deployment with new providers, our priors fall back to the spec sheet; principled cold-start strategies (transfer, meta-learning) are not addressed here. \emph{(iv)} Fairness/safety constraints over provider sets (e.g., regional regulation, content policy): out of scope; the policy currently trusts the affordable set as a black box.
